---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# WebAssembly in R using wasmer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/dirkschumacher/wasmr.svg?branch=master)](https://travis-ci.org/dirkschumacher/wasmr)
<!-- badges: end -->

The goal of `wasmr` is to run [WebAssembly](https://developer.mozilla.org/en-US/docs/WebAssembly/Concepts) code from R using the [wasmer](https://wasmer.io/) runtime.

This is a early version with probably bugs, missing features and not the best code quality. But it already works and I am happy to hear feedback. The goal is that this package evolves into something stable.

The package is mainly written in C++ using the C API of the `wasmer` runtime which is written in `rust`. You therefore need a rust compiler to install the package.

## Installation

``` r
remotes::install_github("dirkschumacher/wasmr")
```

Also make sure to have a very recent rust compiler. The best way to install the rust toolchain is to use [rustup](https://rustup.rs/).

## Example

```{r example}
library(wasmr)
```

```{r}
f <- system.file("examples/sum.wasm", package = "wasmr")
instance <- instantiate(f)
instance$exports$sum(10, 20)
```

```{r}
f <- system.file("examples/hello.wasm", package = "wasmr")
instance <- instantiate(f)
memory_pointer <- instance$exports$hello()
hi <- instance$memory$get_memory_view(memory_pointer)
rawToChar(hi$get(1:11))
```

```{r}
f <- system.file("examples/fib.wasm", package = "wasmr")
instance <- instantiate(f)
instance$exports$fib(20)

fib <- function(n) {
  if (n < 2) return(n)
  fib(n - 1) + fib(n - 2)
}

microbenchmark::microbenchmark(
  instance$exports$fib(20),
  fib(20)
)
```

## Memory

```{r}
f <- system.file("examples/fib.wasm", package = "wasmr")
instance <- instantiate(f)
instance$memory$get_memory_length()

# grow the number of pages
instance$memory$grow(1)
instance$memory$get_memory_length()
```

```{r}
f <- system.file("examples/hello.wasm", package = "wasmr")
instance <- instantiate(f)
memory_pointer <- instance$exports$hello()
memory <- instance$memory$get_memory_view(memory_pointer)
memory$get(1:11)
rawToChar(memory$get(1:11))

# you can also write to the internal memory
memory$set(5:6, charToRaw("o_"))
rawToChar(memory$get(1:11))
```


### Imports

```{r}
set.seed(42)
imports <- list(
  env = list(
    add = typed_function(
      function(a, b) {
        # use R's RNG to add a random number to the result
        a + b * as.integer(runif(1) * 100)
      },
      param_types = c("I32", "I32"),
      return_type = "I32"
    )
  )
)
f <- system.file("examples/sum_import.wasm", package = "wasmr")
instance <- instantiate(f, imports)
# sum: add(a, b) + 42
instance$exports$sum(1, 5)
```

## Caveats and limitations

* It is not feature complete and does not support everything that `wasm` supports
* Imported functions can only have integer parameters
* No globals
* There is hardly any documentation except for the examples
* `I32/I64` are mapped to `IntegerVector` and `F32/F64` to `NumericVector`. Currently no way to differentiate.
* WIP

## Inspiration and References

* [wasmer Python](https://github.com/wasmerio/python-ext-wasm) - haven't tried it but looked at the API/examples
* [wasmer](https://github.com/wasmerio/wasmer) - especially the C api and the tests give some good examples.
* @jeroen 's [rust template](https://github.com/r-rust/hellorust)

## Contribute

While this is work in progress, the best way to contribute is to test the package and write/comment on issues. Before sending a PR it would great to post an issue first to discuss.

## License

MIT just like wasmer.
