---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# WebAssembly in R using wasmer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/dirkschumacher/wasmrrr.svg?branch=master)](https://travis-ci.org/dirkschumacher/wasmrrr)
<!-- badges: end -->

The goal of wasmrrr is to run [WebAssembly](https://developer.mozilla.org/en-US/docs/WebAssembly/Concepts) code from R using the [wasmer](https://wasmer.io/) runtime.

This is a super early version with bugs and missing features. It is a first prototype for me to learn more about WebAssembly, but will evolve hopefully into something stable. It has the appropriate code quality of a prototype :).

## Installation

``` r
remotes::install_github("dirkschumacher/wasmr")
```

Also make sure to have a very recent rust compiler. The best way to install the rust toolchain is to use [rustup](https://rustup.rs/).

## Example

```{r example}
library(wasmr)
```

```{r}
f <- system.file("examples/sum.wasm", package = "wasmr")
instance <- instantiate(f)
instance$exports$sum(10, 20)
```

```{r}
f <- system.file("examples/hello.wasm", package = "wasmr")
instance <- instantiate(f)
memory_pointer <- instance$exports$hello()
(hi <- instance$memory$get_memory_view(memory_pointer))
rawToChar(hi)
```

```{r}
f <- system.file("examples/fib.wasm", package = "wasmr")
instance <- instantiate(f)
instance$exports$fib(20)

fib <- function(n) {
  if (n < 2) return(n)
  fib(n - 1) + fib(n - 2)
}

microbenchmark::microbenchmark(
  instance$exports$fib(20),
  fib(20)
)
```

## Memory

```{r}
f <- system.file("examples/fib.wasm", package = "wasmr")
instance <- instantiate(f)
instance$memory$get_memory_length()

# grow the number of pages
instance$memory$grow(1)
instance$memory$get_memory_length()
```

### Imports

```{r}
set.seed(42)
imports <- list(
  env = list(
    add = typed_function(
      function(a, b) {
        a + b * runif(1) * 100
      },
      param_types = c("I32", "I32"),
      return_type = c("I32")
    )
  )
)
f <- system.file("examples/sum_import.wasm", package = "wasmr")
instance <- instantiate(f, imports)
instance$exports$sum(1, 5)
```


## Inspiration and References

* [wasmer Python](https://github.com/wasmerio/python-ext-wasm) - haven't tried it but looked at the API/examples
* [wasmer](https://github.com/wasmerio/wasmer) - especially the C api and the tests give some good examples.
* @jeroen's [rust template](https://github.com/r-rust/hellorust)

## Contribute

While this is work in progress, the best way to contribute is to test the package and write/comment on issues. Before sending a PR it would great to post an issue first to discuss.

## License

MIT just like wasmer.
