---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# WebAssembly in R using wasmer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/dirkschumacher/wasmrrr.svg?branch=master)](https://travis-ci.org/dirkschumacher/wasmrrr)
<!-- badges: end -->

The goal of wasmrrr is to run [WebAssembly](https://developer.mozilla.org/en-US/docs/WebAssembly/Concepts) code from R using the [wasmer](https://wasmer.io/) runtime.

This is a super early version with bugs and missing features. It is a first prototype for me to learn more about WebAssembly, but will evolve hopefully into something stable. It has the appropriate code quality of a prototype :).

## Installation

``` r
remotes::install_github("dirkschumacher/wasmerrr")
```

## Example

```{r example}
library(wasmrrr)
```

```{r}
f <- system.file("examples/sum.wasm", package = "wasmrrr")
instance <- instantiate(f)
instance$exports$sum(10, 20)
```

```{r}
f <- system.file("examples/hello.wasm", package = "wasmrrr")
instance <- instantiate(f)
memory_pointer <- instance$exports$hello()
(hi <- instance$memory$get_memory_view(memory_pointer))
rawToChar(hi)
```

```{r}
f <- system.file("examples/fib.wasm", package = "wasmrrr")
instance <- instantiate(f)
instance$exports$fib(20)

fib <- function(n) {
  if (n < 2) return(n);
  fib(n - 1) + fib(n - 2);
}

microbenchmark::microbenchmark(
  instance$exports$fib(20),
  fib(20)
)
```


## Inspiration and References

* [wasmer Python](https://github.com/wasmerio/python-ext-wasm) - haven't tried it but looked at the API/examples
* [wasmer](https://github.com/wasmerio/wasmer) - especially the C api and the tests give some good examples.
* @jeroen's [rust template](https://github.com/r-rust/hellorust)
* [Rcpp's modules](http://dirk.eddelbuettel.com/code/rcpp/Rcpp-modules.pdf)

## Todo

* No import support yet - any wasm file with imports does not work
* No Table
* bug fixes
* Read more about the design of wasm 🙈

## License

MIT just like wasmer.
